Gender disparity in computational biology research publications

### 2017-08-04 manuscript.tex
formatting changes

\begin{document}
\vspace*{0.2in}
\linenumbers
% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Gender disparity in computational biology research publications}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Kevin S. Bonham\textsuperscript{1*},
Melanie I. Stefan\textsuperscript{2},
\\
\bigskip
\textbf{1} Microbiology and Immunobiology, Harvard Medical School. Boston, MA, USA
\\
\textbf{2} {Centre for Integrative Physiology, Edinburgh Medical School. Biomedical Sciences, University of Edinburgh. Edinburgh, UK}
\\
\bigskip

@@

@@

When looking at research output in the form of publication and impact, the story is complex: Women tend to publish less than men \cite{Lariviere2013}, are underrepresented in the more prestigious first and last author positions, and publish fewer single-author papers \cite{West2013}. In mathematics, women tend to publish in lower-impact journals \cite{Mihaljevic-Brandt2016}, while in engineering, women publish in journals with higher impact factors \cite{Ghiasi2015}. In general, however, articles authored by women are cited less frequently than articles authoreed by men \cite{Lariviere2013,Ghiasi2016}, which might in part be due to men citing their own work more often than women do \cite{King2016}. Inferring bias in these studies is difficult, since the cause of the disparity between male and female authorship cannot be readily determined. At the same time, when stories of scientific discoveries are told, gender biases are readily identified: Work by female scientists is more likely to be attributed to a male colleague \cite{Rossiter1993}, and biographies of successful female scientists perpetuate gender stereotypes \cite{Fara2013}. Finally, the way in which evidence for gender bias is received is in itself biased:  Male scientists are less likely to accept studies that point to the existence of gender bias than are their female colleagues \cite{Handley2015}.

@@

\begin{figure}[!h]
  \caption{
  A: Mean probability that an author in a given position is female for primary articles indexed in Pubmed with the MeSH term Biology (black) or Computational Biology (grey). The bio dataset is inclusive of papers in the comp dataset. Error bars represent 95\% confidence intervals. B: Mean probability that an author in a given position is female for primary articles indexed in Pubmed with the MeSH term Biology (black) or Computational Biology (grey). The bio dataset is inclusive of papers in the comp dataset. Error bars represent 95\% confidence intervals. C: Mean probability that an author is female for publications in a given year. Error bars represent 95\% confidence intervals. D: Mean probability that the first (F), second (S), penultimate (P) or other (O) author is female for publications where the last author is male ($P_{female}$ < 0.2) or female ($P_{female}$ > 0.8). Papers where the gender of the last author was uncertain or could not be determined were excluded. Error bars represent 95\% confidence intervals.}
  \label{fig1}
\end{figure}

>>>>

\begin{document}
\vspace*{0.2in}
\linenumbers
% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Women are Underrepresented in Computational Biology: an analysis of the scholarly literature in biology, computer science and computational biologys}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Kevin S. Bonham\textsuperscript{1*},
Melanie I. Stefan\textsuperscript{2},
\\
\bigskip
\textbf{1} Microbiology and Immunobiology, Harvard Medical School. Boston, MA, USA
\\
\textbf{2} {Centre for Integrative Physiology, Edinburgh Medical School. Biomedical Sciences, University of Edinburgh. Edinburgh, UK}
\\
\bigskip

@@

\section*{Author Summary}
TODO

@@

When looking at research output in the form of publication and impact, the story is complex: Women tend to publish less than men \cite{Lariviere2013}, are underrepresented in the more prestigious first and last author positions, and publish fewer single-author papers \cite{West2013}. In mathematics, women tend to publish in lower-impact journals \cite{Mihaljevic-Brandt2016}, while in engineering, women publish in journals with higher impact factors \cite{Ghiasi2015}. In general, however, articles authored by women are cited less frequently than articles authoreed by men \cite{Lariviere2013,Ghiasi2015}, which might in part be due to men citing their own work more often than women do \cite{King2016}. Inferring bias in these studies is difficult, since the cause of the disparity between male and female authorship cannot be readily determined. At the same time, when stories of scientific discoveries are told, gender biases are readily identified: Work by female scientists is more likely to be attributed to a male colleague \cite{Rossiter1993}, and biographies of successful female scientists perpetuate gender stereotypes \cite{Fara2013}. Finally, the way in which evidence for gender bias is received is in itself biased:  Male scientists are less likely to accept studies that point to the existence of gender bias than are their female colleagues \cite{Handley2015}.

@@

\begin{figure}[!h]
  \caption{
  A: Mean probability that an author in a given position is female for primary articles indexed in Pubmed with the MeSH term Biology (black) or Computational Biology (grey). The bio dataset is inclusive of papers in the comp dataset. Error bars represent 95\% confidence intervals. B: Mean probability that an author is female for publications in a given year. Error bars represent 95\% confidence intervals. C: Mean probability that the first (F), second (S), penultimate (P) or other (O) author is female for publications where the last author is male ($P_{female}$ < 0.2) or female ($P_{female}$ > 0.8). Papers where the gender of the last author was uncertain or could not be determined were excluded. Error bars represent 95\% confidence intervals.}
  \label{fig1}
\end{figure}
